---
title: "Introduction to tariffTheme"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to tariffTheme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 6,
  dpi = 300
)
```

## Introduction

The `tariffTheme` package provides a consistent visual style for tariff data visualizations. It includes a custom ggplot2 theme, color palettes, and utility functions for adding logos and styling elements.

## Getting Started

First, load the required packages:

```{r setup}
library(ggplot2)
library(data.table)
library(tariffTheme)
```

If you'd like to use the included Merriweather font, you can load it with:

```{r fonts, eval=FALSE}
# Register the Merriweather font (only needed once per R session)
load_tariff_fonts()
```

## Basic Theme Usage

The package includes a built-in dataset of country tariffs that we can use for our examples:

```{r data_exploration}
# Take a look at the first few rows of the dataset
head(country_tariffs)
```

Let's create a basic visualization with the tariff theme:

```{r basic_plot}
# Create a simple bar chart with default ggplot2 theming
basic_plot <- ggplot(country_tariffs[1:10], 
                    aes(x = country, y = reciprocal_tariff_adjusted)) +
  geom_bar(stat = "identity") +
  labs(title = "Reciprocal Tariffs by Country",
       x = "Country",
       y = "Tariff Rate (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the basic plot
print(basic_plot)

# Now apply the tariff theme
themed_plot <- basic_plot +
  theme_tariff() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the themed plot
print(themed_plot)
```

## Using Color Scales

The `tariffTheme` package provides custom color scales that complement the theme:

```{r color_scales}
# Create a plot with color differentiation
color_plot <- ggplot(country_tariffs[1:10], 
                    aes(x = country, y = reciprocal_tariff_adjusted, fill = country)) +
  geom_bar(stat = "identity") +
  labs(title = "Tariffs with Custom Color Scale",
       x = "Country",
       y = "Tariff Rate (%)") +
  theme_tariff() +
  scale_fill_tariff() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

# Display the plot with custom colors
print(color_plot)
```

## Adding a Logo and Borders

One of the key features of the `tariffTheme` package is the ability to add a logo and styled borders to your plots:

```{r logo_example}
# Create a basic plot with the tariff theme
logo_plot <- ggplot(country_tariffs[1:5], 
                   aes(x = country, y = reciprocal_tariff_adjusted)) +
  geom_bar(stat = "identity", fill = yellow_color) +
  geom_text(aes(label = paste0(reciprocal_tariff_adjusted, "%")), 
            position = position_stack(vjust = 0.5),
            color = "black") +
  labs(title = "Tariffs with Logo and Borders",
       x = "Country",
       y = "Tariff Rate (%)") +
  theme_tariff() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add the logo and borders
final_plot <- add_logo(logo_plot)

# Display the final plot
print(final_plot)
```

## Styling Labels

The package also provides helper functions for styling labels and annotations:

```{r label_styling}
# Create a plot with labels
label_plot <- ggplot(country_tariffs[1:5], 
                    aes(x = country, y = reciprocal_tariff_adjusted)) +
  geom_bar(stat = "identity", fill = light_blue_color) +
  geom_label(aes(label = paste0(reciprocal_tariff_adjusted, "%")), 
             position = position_stack(vjust = 0.5)) +
  labs(title = "Tariffs with Styled Labels",
       x = "Country",
       y = "Tariff Rate (%)") +
  theme_tariff() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Apply label styling
styled_label_plot <- style_tariff_label(label_plot)

# Display the plot with styled labels
print(styled_label_plot)
```

## Example: Random Sampling for Better Visualization

Sometimes we might want to work with a subset of the data for better visualization. Here's how we can create a plot with a random sample of countries:

```{r random_sample}
# Set seed for reproducibility
set.seed(42)

# Sample 10 random countries
tariff_sample <- country_tariffs[sample(.N, 10)]

# Make country a factor variable with randomized levels for better display
tariff_sample[, country := factor(country, levels = sample(country))]

# Create a bar chart with the sample
sample_plot <- ggplot(tariff_sample, aes(x = country, y = reciprocal_tariff_adjusted)) +
  geom_bar(stat = "identity", width = 0.7, fill = yellow_color) +
  geom_text(aes(label = paste0(reciprocal_tariff_adjusted, "%")), 
            position = position_stack(vjust = 0.5),
            color = "black") +
  labs(
    title = "Reciprocal Tariffs",
    x = "Country",
    y = "U.S.A Discounted Reciprocal Tariffs"
  ) +
  theme_tariff() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Add logo and borders
final_sample_plot <- add_logo(sample_plot, corner_radius = 0.03)

# Display the final plot
print(final_sample_plot)
```

## Conclusion

The `tariffTheme` package provides a consistent and visually appealing way to create tariff data visualizations. By combining the custom theme, color scales, and utility functions, you can quickly create professional-looking plots that maintain a consistent brand identity.